### 2018-06-05 latex/egsr17-cgf/response/review_response.tex

\documentclass[11pt,letterpaper,headsepline,DIV15]{scrartcl}

\def\bf{\normalfont\bfseries}

\usepackage{enumitem}
\usepackage{bm} % for bold greek symbols
\usepackage{color}

\usepackage{alltt}
%\renewcommand{\ttdefault}{txtt}

%checkmark
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\color{green}\ding{51}\color{black}}%
\newcommand{\xmark}{\color{red}\ding{55}\color{black}}%

% strike out text
\usepackage{soul,xcolor}

\definecolor{red}{RGB}{120,0,0}
\definecolor{gray}{RGB}{150,150,150}
\definecolor{todoColor}{RGB}{120,30,60}

\def\todo#1{{\textbf{\em \textcolor{todoColor}{\{TODO: #1\}}}}}
\def\todogm#1{{\textbf{\em \textcolor{todoColor}{\{TODO-G: #1\}}}}}
\def\todofs#1{{\textbf{\em \textcolor{todoColor}{\{TODO-F: #1\}}}}}

\definecolor{answerColor}{RGB}{40,60,140}
%\newcounter{aCounter}
%\setcounter{aCounter}{0}
%%\newcommand{\answer}[1]{\textcolor{answerColor}{\textbf{#1}}}
%%\newcommand{\answer}[0]{\noindent\stepcounter{aCounter}\textcolor{answerColor}{Reply	 \arabic{aCounter} }}
%\newcommand{\answer}[1]{\noindent\stepcounter{aCounter}\textcolor{answerColor}{\bf Reply \arabic{aCounter}\\}{\textcolor{answerColor}{#1}}}
%%\newcommand{\answer}[2]{\noindent\stepcounter{aCounter}\textcolor{answerColor}{\label{#1}\bf Reply \arabic{aCounter}\\}{\textcolor{answerColor}{#2}}}

% -------- comment --------------
\usepackage{verbatim}
\newcounter{cmntcounter}
\renewcommand{\thecmntcounter}{\arabic{cmntcounter}}

\newcommand{\nextcomment}[1]{\refstepcounter{cmntcounter}\label{#1}\textcolor{red}{\textbf{comment-\thecmntcounter}}}

% -------- reply --------------
\newcounter{rplcounter}
%\renewcommand{\thefpcounter}{\thechapter.\arabic{fpcounter}}
%\renewcommand{\thefpcounter}{\thesection.\arabic{fpcounter}}
\renewcommand{\therplcounter}{\arabic{rplcounter}}

\newenvironment{reply}[2]{%
\refstepcounter{rplcounter}%
\label{#1}%
\noindent\textbf{\textcolor{answerColor}{Reply~\therplcounter}\\}%
\color{answerColor}
}%
{}%

\newcommand{\tocomment}[1]{\bf#1}

\usepackage[T1]{fontenc}
\newcommand{\changefont}[3]{
\fontfamily{#1} \fontseries{#2} \fontshape{#3} \selectfont}
\changefont{phv}{m}{n}
\renewcommand{\familydefault}{\sfdefault}

%\title{{\small Cover Letter for VisWeek 2013 -- Paper Id 130, Minor Revision}\\

% {\Large Coronal Holes Extraction Based on Isocline Surfaces}}

\title{\normalsize Cover Letter for EGSR2018-- Submission 1027 (EI\&I), Revised Version}

\vspace{1cm}
%\author{M. \"Uffinger, F. Sadlo, T. Ertl}
%\date{2013-11-22}
\date{}

\begin{document}

\maketitle
\noindent Dear Program Chairs, Dear Reviewers,
\\

\noindent thank you for reviewing our submission and accepting it for publication at EGSR2018 EI\&I track with a minor revision.
\\

\noindent Common among the reviews is the request for more rendering evaluations, which unfortunately we have not available yet. While we took care in presenting results, which validated the theory (pointsource problem) and demonstrated applicability in rendering (Nebulae), we agree that more results are needed. The lack thereof is one of the reasons, why we submitted to the EI\&I track. We are planning to have more renderings of different datasets for the next stage of this project.
\\

\noindent In the following, we cite the original comments with non-proportional black font and use bold font for the parts we address in the subsequent replies, which are in blue color. Attached to this response, is the revised version where our changes in the text are highlighted in red.
\\
\\
\noindent Respectfully,\\
the Authors.

\clearpage

%\section*{Improved TVCG submission}

\vspace{0.5cm}

{\scriptsize\begin{alltt}
-------------------- Review 17707 --------------------

Although finite-element-based solutions to the RTE are well known in many areas, this topic is underexplored in
computer graphics. The solution proposed in the paper seems practical, and the paper has the potential to encourage
future research along this direction.

\tocomment{
The main issue of this paper is the lack of rendering evaluations: the only actual rendered results are in Figure 1, and the P_5 result is only slightly better than classical diffusion approximation and arguably worse than flux-limited diffusion. Thus, the practical usefulness of the proposed technique for graphics applications remain unclear.
}
\end{alltt}}

\begin{reply}{}{}
We agree with the reviewer on that the practical usefulness of our technique is unclear. In fact, at this stage, we admit that in comparison to flux-limited diffusion, our technique does not appear to be competitive. We mention this in the conclusion of our paper. At this point, the value in our submission is, that it closes a knowledge gap in the graphics community on the PN-method, presents the real-valued PN-equations and also introduces an interesting approach for solving general coupled PDE's. Making this method competitive in comparison to other deterministic approaches is our plan going forward.
\end{reply}

{\scriptsize\begin{alltt}
-------------------- Review 17724 --------------------
This paper is huge in terms of scope, and it opens exploration in multiple ways, even with very few results.
This paper is a nice contribution to the E&I Track.

There are some open questions that might need some discussion:
\tocomment{
- The need to create a custom CAS should be really justified, as it seems like overkill. There are some CAS libraries (SymPy for Python,
GiNaC for C++) that would probably fit perfectly this approach, and that are able to generate C++ code and LaTeX. Both of them enable exploration and parsing of their "tree". In any case, for a potential future version it might make sense to opensource the specific CAS library, write a technical report about it and cite it. Right now it adds some extra noise to the paper's story (which might be summarized as "we use a CAS library to derive and generate stencil code for the P-N equations").}
\end{alltt}}

\begin{reply}{}{}
We agree with the reviewer in that other off-the-shelve computer algebra frameworks, such as SymPy, would be suitable for the tasks of expression manipulation and automated discretization. The reason why we went for our own solution is simply, that we started with a very small lightweight implementation for initial tests and stuck with it throughout the project.
\newline
\newline
The reviewer suggests, that simply pointing to an existing computer algebra framework would allow to summarize things as "we use a CAS library to derive and generate stencil code for the P-N equations". While we agree, that the fact that we created our own computer algebra representation is of no importance to the paper, we find that covering how to use any of such a representation for automated discretization is an important aspect of our work. Computer algebra frameworks, such as SimPy, are primarily known for and used to perform symbolic computation. We feel, that it is worth covering in more detail, how such a framework can be used for automated discretization. This argument is backed up by the fact, that none of the off-the-shelf packages offers a manipulation step for discretizing a given equation. We find that implementing such a manipulation step is not trivial, especially in the presence of coupling and staggered grid locations.
\newline
\newline
So in order to account for the reviewers comment, we made changes to the text in various places, which de-emphasize the use of our own implementation.
%\begin{itemize}
%	\item abstract: We intoduce a computer algebra framework $\rightarrow$ We use a computer algebra framework
%\end{itemize}
Further, we reworked section 4, which was used to introduce our own CAS-framework. We now state, that any computer algebra representation can be used and reference SymPy as an example. We also shortened this section by removing the text which was dedicated to explaining the capabilities of our representation as those are only a subset of what most of the alternatives can do. We kept a paragraph, which summarizes the concept of such frameworks and also kept figure 4, which expresses the idea of computer algebra representations visually. Arguably, this is known by most of the readers, but we find it useful to briefly summarize the concept, because we refer to it in more detail in section 6.1, when we explain the automated discretization.
\newline
\newline
Section 6.1 explains the implementation of a discretization step, independent from the computer algebra framework used. As already mentioned, such a manipulation would have to be implemented for any framework and this section gives a good summary on how to go about it. We therefore see enough value in this section to keep it as is.
\end{reply}

{\scriptsize\begin{alltt}
\tocomment{
- The use of P-N for diffusion in participating media might be an unviable solution. The paper in the discussion section points to avenues of future work in terms of speed-up (multigrid linear solver, for instance) but there should be some avenue of future work regarding convergence. Results look adequate, but would it ever converge, given enough order in spherical harmonics? In which cases the solution is closer to the ground truth? Which kinds of participating media are better represented? From my point of view that is much more important than render time.}
\end{alltt}}

\begin{reply}{}{}
\label{rpl:gg}
We very much agree with this comment. To be clear: The $P_N$-method is only not converging, when vacuum or close-to vacuum regions are present, because the matrix $A$ is singular in these cases. If the extinction coefficient is sufficiently large throughout the domain, the method does converge (but might still require a lot of computation for that).
\newline
\newline
The presence of vacuum regions and its problem with conergence was particularily important to us, as vacuum regions are common in rendering. This problem seems to be closely related to second order methods, which we mention in the conclusion section of the paper (e.g. the work of Hansen et al.~\cite{Hansen14}). Some of those second order methods are said to have the benefit of being able to deal with vacuum regions. They are an active research topic in the neutron transport community. By solving the normal form ($A^TA$), we basically solved the naive second order form of the RTE and we found that it is able to deal with vacuum regions in the sense, that it does not break down formally (e.g. due to division by zero), but at the same time it does not converge, since $A^TA$ is still singular if $A$ happens to be singular. A paragraph in the conclusion section is dedicated to this insight.
\newline
\newline
We have not been very clear about the singularity of $A$ in the presence of vacuum. We added a comment on this in section 6.2.
\newline
\newline
There are different second order forms, such as Even-Odd-Parity (EOP), Self-adjoint angular flux (SAAF) or the least-squares form introduced by Hansen et al.~\cite{Hansen14} and evaluating their properties for rendering applications would be a very interesting avenue for future work. We mention this at the end of the conclusion section. However, these research questions really do exceed the scope of our paper, which is primarily about giving the theory and a practical implementation of the $P_N$-mehod with some validation.
\end{reply}

{\scriptsize\begin{alltt}
\tocomment{
- The linear system to be solved during runtime targets the normal form A^T A because it is not simmetric. However, the used library (Eigen) provides a variety of solvers that would probably ensure accuracy even for non-simmetric matrices (LU? SVD? QR?).
In that case, why is it better to use the normal form with a naturally higher condition number?}
\end{alltt}}

\begin{reply}{}{}
If no vacuum regions are present, the mentioned alternatives would work for both, the first order form ($A\vec{x}=\vec{b}$) and the normal form of the problem. The reason why we introduced the normal form is, that it enables the use of standard iterative methods, such as conjugate gradient. These iterative methods allow balancing the computation time against accuracy by tweaking the convergence threshold. We adjusted the text in section 6.2 to outline this decision.

As mentioned in reply~\ref{rpl:gg}, we have not been explicit about the singularity of $A$ in the presence of vacuum. In these cases, iterative solvers will not converge and solvers, such as LU factorization will not be applicable. We added a remark in the text in section 6.2.
%mention inverse of the scattering/collision operator
%introduce text referencing the lspn stuff and talking about the relation to normal form of the RTE

%When we take the normal form of the linear system $A\vec{x}=Q$ (which we arrive at after discretization), we essentially get the coefficient matrix of the naive second order form of the RTE. The neutron transport literature states, that standard first order forms are not able to cope with vacuum regions, since their solution requires inversion of the scattering/collision operator. And those break down in vacuum. We saw that solvers, such as LU were not able to converge. such as the least-squares $P_N$ method introduced by 
\end{reply}

{\scriptsize\begin{alltt}
\tocomment{
Last, I would like to congratulate the authors for their work and I would like to encourage them to keep this research until a final version, that most surely I would like to read.}
\end{alltt}}

\begin{reply}{}{}
We are really thankful for your comment and encouragement. Deterministic methods are a niche topic and it is unclear, whether they will find a place in rendering long-term. However, investigating these methods and providing a clear understanding about their traits and their applicability is valuable research for which the EI\&I track seems to be a good place. We are very happy that all reviewers seem to agree.
\end{reply}

{\scriptsize\begin{alltt}
-------------------- Review 17681 --------------------

The previous paragraphs explain the reasons why I will reference the paper. I think that beyond the CG community,
this paper can have a larger influence in many work dealing with the volume Rendering.

It is very important to make sure that the supplemental materials and the source code are released at the same time
than the paper.

\tocomment{Note to the authors: the section 3.1.2 of the supplemental material is empty. Make sure to correct for the final version of the paper.}
\end{alltt}}
\begin{reply}{}{}
We completed section 3.1.2 in the supplemental material on the real-valued spherical harmonics expansion of the phase function and also fixed some minor notation issues throughout the text.
\end{reply}

%\bibliographystyle{plain}
\bibliographystyle{eg-alpha}
\bibliography{bibliography}

\end{document}

### 2018-06-08 latex/egsr17-cgf/introduction.tex
revision 2

\section{Introduction}

Simulating light transport in paticipating media remains a challenging problem for image synthesis in computer graphics. Due to their ability to produce unbiased results and conceptional simplicity, Monte Carlo based techniques have become the standard approach. The main downside of these methods are their computational demands when rendering participating media, which exhibits strong scattering or anisotropy.

@@

The work on path guiding techniques from recent years has shown, how approximate representations of the steady-state transport in a scene can be used to accelerate Monte Carlo integration techniques, such as path tracing. Instead of generating these approximate representations using Monte Carlo methods, deterministic methods may offer a viable alternative. Hybrid methods could combine benefits of deterministic methods while generating unbiased results.

Deterministic methods also lend themselves to applications where fast approximate solutions are preferable over correct, but slowly converging results. For this reason, we find it important to study deterministic methods and have a solid understanding about their characteristics and performance traits for typical rendering problems.

The $P_N$-method is a deterministic method, which is popular in fields like medical imaging and nuclear sciences, but has not found use in computer graphics thus far. The purpose and main contribution of our paper is to gain a solid understanding of its foundations and present a method for using it in the context of rendering. In particular, we present these theoretical and practical contributions:
\begin{itemize}
	\item We derive and present the time-independent real-valued $P_N$-equations. To our knowledge, we are the first to derive and write down their compact form.
	\item We introduce a staggered-grid solver, for which we generate stencil code automatically from a computer algebra representation of the $P_N$-equations. This allows us to deal with the increasingly complex equations, which the $P_N$-method produces for higher order. It further allows our solver to be used for any (potentially coupled) partial differential equation, which results in a system of linear equations after discretization.
	\item Finally, we compare the $P_N$-method for higher orders against flux-limited diffusion and ground truth Monte Carlo integration.
\end{itemize}

>>>>

\section{Introduction}
\vspace{0.1in}

Simulating light transport in participating media remains a challenging problem for image synthesis in computer graphics. Due to their ability to produce unbiased results and conceptual simplicity, Monte Carlo based techniques have become the standard approach. The main downside of these methods are their computational demands when rendering media with strong scattering or anisotropy.

@@

The work on path guiding techniques from recent years has shown how approximate representations of the steady-state transport in a scene can be used to accelerate Monte Carlo integration techniques, such as path tracing. Instead of generating these approximate representations using Monte Carlo methods, deterministic methods may offer a viable alternative. Hybrid methods could combine the performance benefits of deterministic methods with accurate and unbiased Monte Carlo results.

Deterministic methods also lend themselves to applications where fast approximate solutions are preferable over correct, but slowly converging results. For this reason, we suggest it is important for volume-rendering researchers to study deterministic methods and have a solid understanding of their characteristics and performance traits for typical rendering problems.

The $P_N$-method is a deterministic method which is popular in fields like medical imaging and nuclear sciences, but has not found use in computer graphics thus far. The purpose and main contribution of our paper is to gain a solid understanding of its foundations and present a method for using it in the context of rendering. In particular, we present these theoretical and practical contributions:
\begin{itemize}
	\item We derive and present the time-independent real-valued $P_N$-equations. To our knowledge, we are the first to derive and write down their compact form.
	\item We introduce a staggered-grid solver, for which we generate stencil code automatically from a computer algebra representation of the $P_N$-equations. This allows us to deal with the increasingly complex equations, which the $P_N$-method produces for higher order. It further allows our solver to be used for any (potentially coupled) partial differential equation, which results in a system of linear equations after discretization.
	\item Finally, we compare the $P_N$-method for higher orders against flux-limited diffusion and ground truth Monte Carlo integration.
\end{itemize}

### 2018-06-08 latex/egsr17-cgf/previous_work.tex
revision 2

Light transport in participating media is governed by the radiative transfer equation~(RTE), first studied in the context of astrophysics by Chandrashekar~\cite{Chandrasekhar60} and later introduced to computer graphics by Kajiya~\cite{Kajiya86}. Today, this equation is typically solved using Monte Carlo methods. In strongly scattering media or in the presence of highly anisotropic phase functions, these methods can become prohibitively expensive. Consider a high albedo media like milk, where tracing paths with many scattering events is necessary.

In contrast to path-tracing, the $P_N$-method gives a solution by solving a system of linear equations. It is derived by discretizing the angular variable of the radiative transfer equation into spherical harmonics (SH). This gives rise to a set of coupled, complex-valued partial differential equations, called $P_N$-equations. The subscript $N$ refers to the truncation order of the spherical harmonics.

The $P_N$-method has a long history in other fields and was never applied in graphics. Kajia~\cite{Kajiya84} explained the theory, but did not give any details on implementation or how to solve it. In fact, as Max~\cite{Max95} pointed out, it is not clear if Kajiya succeeded at all at applying the method, as all of the results in his paper were produced with a simpler method. This is further strengthened by our observation, that a straight forward finite difference discretization of the $P_N$-equations produces unuseable results, due to oscillation artifacts in the solution. In our paper, we derive the real-valued $P_N$-equations along with a staggered-grid solver, that produces artifact free solutions.

@@

VEF is based on the observation, that the behavior of volumetric transport is closely linked to the density of the material: In the diffusive regime, thick highly scattering media is present and causes photons to change directions in short succession, while the transport regime in thin and highly absorbing media causes photons to travel long straight lines. In the absence of scattering (e.g. pure vacuum), the non-linear diffusion coefficient turns the diffusion equation into an advection equation. The idea behind VEF is to derive a better diffusion coefficient by seeing it as a mean to spatially blend between the diffusive and pure transport regime.

It is an open and unresolved question, whether the $P_N$-method with truncation at higher order, gives any benefit over using first order non-linear diffusion. This question is also been raised in other domains~\cite{Olson00} and resolving it is one of our motivations.

>>>>

Light transport in participating media is governed by the radiative transfer equation~(RTE), first studied in the context of astrophysics by Chandrashekar~\cite{Chandrasekhar60} and later introduced to computer graphics by Kajiya~\cite{Kajiya86}. Today, this equation is typically solved using Monte Carlo methods. In strongly scattering media or in the presence of highly anisotropic phase functions, these methods can become prohibitively expensive. Consider a high albedo medium like milk, where tracing paths with many scattering events is necessary.

In contrast to path-tracing, the $P_N$-method gives a solution by solving a system of linear equations. It is derived by discretizing the angular variable of the radiative transfer equation into spherical harmonics (SH). This gives rise to a set of coupled, complex-valued partial differential equations, called the $P_N$-equations. The subscript $N$ refers to the spherical harmonics truncation order.

The $P_N$-method has a long history in other fields and was never applied in graphics. Kajia~\cite{Kajiya84} explained the theory, but did not give any details on implementation or how to solve it. In fact, as Max~\cite{Max95} pointed out, it is not clear if Kajiya succeeded at all at applying the method, as all of the results in his paper were produced with a simpler method. This is further strengthened by our observation, that a straight forward finite difference discretization of the $P_N$-equations produces unuseable results, due to oscillation artifacts in the solution. We derive the real-valued $P_N$-equations with a staggered-grid solver, that produces artifact free solutions.

@@

VEF is based on the observation, that the behavior of volumetric transport is closely linked to the density of the material: In the diffusive regime, thick highly scattering media is present and causes photons to change directions in short succession, while the transport regime in thin and highly absorbing media causes photons to travel long straight lines. In the absence of scattering (e.g. pure vacuum), the non-linear diffusion coefficient turns the diffusion equation into an advection equation. The idea behind VEF is to derive a better diffusion coefficient by seeing it as a means to spatially blend between the diffusive and pure transport regime.

It is an open and unresolved question, whether the $P_N$-method with truncation at higher order, gives any benefit over using first order non-linear diffusion. This question has also been raised in other domains~\cite{Olson00} and resolving it is one of our motivations.

### 2018-06-08 latex/egsr17-cgf/rendering.tex
revision 2

We use an approach similar to Koerner et al.~\cite{Koerner14}, where we seperate the radiance field into single scattered light $L_{ss}$ and multiple scattered light $L_{ms}$:
\begin{align}
L\left(\vec{x},\omega\right) = L_{ss}\left(\vec{x},\omega\right) + L_{ms}\left(\vec{x},\omega\right)
\end{align}

The single scattered light is folded into the emission term $Q$:
\begin{align}
Q(\vec{x}, \omega) = L_{ss}(\vec{x}, \omega) = \sigma_s\left(\vec{x}\right)\int_\Omega{ p\left(\omega'\rightarrow\omega\right) L_{u}\left(\vec{x}, \omega'\right)\ud\omega' }
\end{align}

This means that, our solver will solve for the multiple scattered light $L_{ms}$. The quantity $L_u$ is the uncollided light, which was emitted from the light source and attenuated by the volume without undergoing any scattering event.
We compute it using a few light samples per voxel, which quickly converge to a useful result for Dirac Delta light sources.

@@

For rendering, we use a simple forward path tracing approach, where we start tracing from the camera. At the first scattering event, we use next event estimation to account for $L_{ss}$. Then we sample a new direction according to the phase function. Instead of continuing tracing into the new direction, we evaluate the inscattering integral using $\widehat{L}_{ms}$. The SH coefficients at $\vec{x}$ are found by trilinear interpolation from the voxel grid of SH coefficients.

>>>>

We use an approach similar to Koerner et al.~\cite{Koerner14}, where we seperate the radiance field into single scattered light $L_{ss}$ and multiple scattered light $L_{ms}$:
\begin{align}
L\left(\vec{x},\omega\right) = L_{ss}\left(\vec{x},\omega\right) + L_{ms}\left(\vec{x},\omega\right)
\ .
\end{align}

The single scattered light is folded into the emission term $Q$:
\begin{align}
Q(\vec{x}, \omega) = L_{ss}(\vec{x}, \omega) = \sigma_s\left(\vec{x}\right)\int_\Omega{ p\left(\omega'\rightarrow\omega\right) L_{u}\left(\vec{x}, \omega'\right)\ud\omega' }
\ .
\end{align}

This means that, our solver will solve for the multiple scattered light $L_{ms}$. The quantity $L_u$ is the uncollided light, which was emitted from the light source and attenuated by the volume without undergoing any scattering event.
We compute it using a few light samples per voxel, which quickly converge to a useful result for Dirac delta light sources.

@@

For rendering, we use a simple forward path tracing approach, where we start tracing from the camera. At the first scattering event, we use next event estimation to account for $L_{ss}$. Then we sample a new direction according to the phase function. Instead of continuing tracing into the new direction, we evaluate the in-scattering integral using $\widehat{L}_{ms}$. The SH coefficients at $\vec{x}$ are found by trilinear interpolation from the voxel grid of SH coefficients.

### 2018-06-08 latex/egsr17-cgf/section_conclusion.tex
revision 2

The literature in other fields often states, that the $P_N$ method---when solved in normal form like we do---is able to deal with vacuum regions. We found this misleading. The $P_N$-method in normal form indeed does not break down completely in the presence of vacuum like diffusion based methods do (due to $\sigma_t^{-1}$ in the diffusion coefficient). However, in the presence of vacuum, the condition number of the system matrix becomes infinite and does not converge either. Therefore $P_N$ based methods also require minimum thresholding of the extinction coefficient and offer no benefit for vacuum regions.

Much more work needs to be done in order to make our method competitive to alternative solutions. We believe this can be made possible by employing a multigrid scheme for solving the linear system of equations. We implemented a multigrid solver, but did not find the expected performance improvements. This is possibly due to the coupling between coefficients within a voxel, which does not work well together with the smoothing step. We want to study this further in the future.

Unique to our system is, that it uses a computer algebra representation of the equation to solve as input. Discretization in angular and spatial domain is done using manipulation passes on the representation. The stencil code, which is used to propagate the system of linear equations, is generated from the expression tree. This way, we can easily deal with coupled PDE's and avoid the time consuming and error prone process of writing stencil code by hand.

>>>>

\vspace{0.5in}

The literature in other fields often states, that the $P_N$ method---when solved in normal form like we do---is able to deal with vacuum regions. We found this misleading. The $P_N$-method in normal form indeed does not break down completely in the presence of vacuum as diffusion based methods do (due to $\sigma_t^{-1}$ in the diffusion coefficient). However, in the presence of vacuum, the condition number of the system matrix becomes infinite and does not converge either. Therefore $P_N$ based methods also require minimum thresholding of the extinction coefficient and offer no benefit for vacuum regions.

Much more work needs to be done in order to make the $P_N$ method competitive in performance to alternative solutions for volume rendering. We believe this can be made possible by employing a multigrid scheme for solving the linear system of equations. We implemented a multigrid solver, but did not find the expected performance improvements. This is possibly due to the coupling between coefficients within a voxel, which does not work well together with the smoothing step. We want to study this in the future.

Unique to our system is that it uses a computer algebra representation of the equation to solve as input. Discretization in angular and spatial domain is done using manipulation passes on the representation. The stencil code, which is used to populate the system of linear equations, is generated from the expression tree. This way, we can easily deal with coupled PDE's and avoid the time consuming and error prone process of writing stencil code by hand.

### 2018-06-08 latex/egsr17-cgf/section_discretized_rte.tex
minor revision
revision 2

Our method derives from the radiative transfer equation (RTE), which expresses the change of the radiance field $L$, with respect to an infinitesimal change of position into direction $\omega$ at point $\vec{x}$:
\begin{align}
%\label{eq:rte}
\left(\nabla\cdot\omega\right)L\left(\vec{x}, \omega \right)
=&
-\sigma_t\left(\vec{x}\right) L\left(\vec{x}, \omega \right)\nonumber\\
&
+\sigma_s\left(\vec{x}\right) \int_{\Omega}
{
p\left(\omega'\cdot\omega\right)L\left(\vec{x}, \omega \right)\ud\omega'
}\nonumber\\
&
+Q\left(\vec{x}, \omega\right)\nonumber
\label{eq:rte}
\end{align}

The left hand side (LHS) is the transport term, and we refer to the terms on the right hand side (RHS) as collision, scattering, and source term, respectively. The symbols $\sigma_t$, $\sigma_s$, $p$, and $Q$ refer to the extinction coefficient, scattering coefficient, phase function and emission term.

The RTE is often given in operator notation, where transport, collision, and scattering are expressed as operators $\mathcal{T}$, $\mathcal{C}$ and $\mathcal{S}$, which are applied to the radiance field $L$:
\begin{align}
\mathcal{T}\left(L\right) = -\mathcal{C}\left(L\right) + \mathcal{S}\left(L\right) + Q
\end{align}

Deterministic methods, like ours, are derived by discretizing the angular and spatial domain. This gives rise to a linear system of equations, which can be solved using standard methods. For the $P_N$-method, the angular variable is first discretized, using a truncated spherical harmonics expansion. This results in the $P_N$-equations, a system of coupled PDE's that still depend on a continuous spatial variable.

@@

Since the radiance field $L$ is real, we use the real-valued SH basis functions $\SHBR^{l,m}$, which are defined in terms of the complex-valued SH basis functions $\SHBC^{l,m}$:
\begin{align}
\SHBR^{l,m}=
\left\{
\begin{array}{lr}
\frac{\iu}{\sqrt{2}}\left(\SHBC^{l,m}-\left(-1\right)^m\SHBC^{l,-m}\right), & \text{for } m < 0\\
\SHBC^{l,m}, & \text{for } m = 0\\
\frac{1}{\sqrt{2}}\left(\SHBC^{l,-m}-\left(-1\right)^m\SHBC^{l,m}\right), & \text{for } m > 0
\end{array}
\right.
\label{eq:sh_real_basis}
\end{align}

We express the projection into the spherical harmonics basis functions with a projection operator $\mathcal{P}$:
\begin{align}
\mathcal{P}^{l, m}(f) = \int_{\Omega}f(\mathbf{x}, \omega) \SHBR^{l,m}(\omega)\,\mathrm{d}\omega = f^{l,m}
\nonumber
\end{align}

The $P_N$-equations are derived by first expressing all direction-dependent parameters in spherical harmonics. The radiance field $L$ in Equation~\ref{eq:rte} is therefore replaced by its SH reconstruction $\widehat{L}$, introducing an error due to truncation at order $N$:
\begin{align}
\widehat{L}\left(\vec{x}, \omega\right) =
\sum_{l=0}^{N}
{
\sum_{m=-l}^{l}
{
L^{l,m}\left(\vec{x}\right)\SHBR^{l,m}\left(\omega\right)
}
}
\approx
L\left(\vec{x}, \omega\right)
\nonumber
\end{align}

After substitution, all angular parameters are expressed in terms of spherical harmonics, but they still depend on the continuous angular variable $\omega$. As a next step, we project each term of the RTE into spherical harmonics, using the projection operator $\mathcal{P}$. This produces a single equation for each $l,m$-pair. The $P_N$-equations therefore can be written as:
\begin{align}
\mathcal{P}^{l,m}\mathcal{T}\left(\widehat{L}\right)
=
-\mathcal{P}^{l,m}\mathcal{C}\left(\widehat{L}\right) 
+\mathcal{P}^{l,m}\mathcal{S}\left(\widehat{L}\right)
+\mathcal{P}^{l,m}\left(Q\right)
\label{eq:pn_operator_notation}
\end{align}

Once the $P_N$-equation have been found, the spatial variable $\vec{x}$ is discretized using a finite difference (FD) voxel grid (using central differences for differential operators).

Following this discretization, the radiance field $L$, is represented as a set of SH coefficients per voxel. Flattening these over all voxels into a single vector, gives the solution vector $\vec{u}$. The RHS vector $\vec{Q}$ is produced similarly. The projected operators can be expressed as linear transformations, which can be collapsed into a single coefficient matrix $A$ (see figure~\ref{fig:matrix_layout}):
\begin{align}
(T+C-S)\vec{u} = A\vec{u} = \vec{Q}
\end{align}

@@

So far, we have only given the $P_N$-equations in high-level operator notation (Equation~\ref{eq:pn_operator_notation}). Carrying out the full derivation creates large, unwieldy equations and requires a string of expansions, manipulations and applications of identities. These are challenging and error-prone if done by hand. We therefore developed a computer algebra representation, which allowed us to derive the the $P_N$-equation in a semi-automatic fashion.

>>>>

Our method derives from the radiative transfer equation (RTE), which expresses the change of the radiance field $L$, with respect to an infinitesimal change of position into direction $\omega$ at point $\vec{x}$:
\begin{align}
%\label{eq:rte}
\left(\nabla\cdot\omega\right)L\left(\vec{x}, \omega \right)
=&
-\sigma_t\left(\vec{x}\right) L\left(\vec{x}, \omega \right)\nonumber\\
&
+\sigma_s\left(\vec{x}\right) \int_{\Omega}
{
p\left(\omega'\cdot\omega\right)L\left(\vec{x}, \omega \right)\ud\omega'
}\nonumber\\
&
+Q\left(\vec{x}, \omega\right)\nonumber
\quad .
\label{eq:rte}
\end{align}

The left hand side (LHS) is the transport term, and we refer to the terms on the right hand side (RHS) as collision, scattering, and source term, respectively. The symbols $\sigma_t$, $\sigma_s$, $p$, and $Q$ refer to the extinction- and scattering coefficient, phase function and emission.

The RTE is often given in operator notation, where transport, collision, and scattering are expressed as operators $\mathcal{T}$, $\mathcal{C}$ and $\mathcal{S}$, which are applied to the radiance field $L$:
\begin{align}
\mathcal{T}\left(L\right) = -\mathcal{C}\left(L\right) + \mathcal{S}\left(L\right) + Q
\quad .
\end{align}

Deterministic methods are derived by discretizing the angular and spatial domain. This gives rise to a linear system of equations, which can be solved using standard methods. For the $P_N$-method, the angular variable is first discretized, using a truncated spherical harmonics expansion. This results in the $P_N$-equations, a system of coupled PDE's that still depend on a continuous spatial variable.

@@

Since the radiance field $L$ is real, we use the real-valued SH basis functions $\SHBR^{l,m}$, which are defined in terms of the complex-valued SH basis functions $\SHBC^{l,m}$:
\begin{align}
\SHBR^{l,m}=
\left\{
\begin{array}{lr}
\frac{\iu}{\sqrt{2}}\left(\SHBC^{l,m}-\left(-1\right)^m\SHBC^{l,-m}\right), & \text{for } m < 0\\
\SHBC^{l,m}, & \text{for } m = 0\\
\frac{1}{\sqrt{2}}\left(\SHBC^{l,-m}-\left(-1\right)^m\SHBC^{l,m}\right), & \text{for } m > 0
\end{array}
\right.
\ .
\label{eq:sh_real_basis}
\end{align}

We express the projection into the spherical harmonics basis functions with a projection operator $\mathcal{P}$:
\begin{align}
\mathcal{P}^{l, m}(f) = \int_{\Omega}f(\vec{x}, \omega) \SHBR^{l,m}(\omega)\,\mathrm{d}\omega = f^{l,m}\left(\vec{x}\right)
\ .
\nonumber
\end{align}

The $P_N$-equations are derived by first expressing all direction-dependent parameters in spherical harmonics. The radiance field $L$ in Equation~\ref{eq:rte} is therefore replaced by its SH reconstruction $\widehat{L}$, introducing an error due to truncation at order $N$:
\begin{align}
\widehat{L}\left(\vec{x}, \omega\right) =
\sum_{l=0}^{N}
{
\sum_{m=-l}^{l}
{
L^{l,m}\left(\vec{x}\right)\SHBR^{l,m}\left(\omega\right)
}
}
\approx
L\left(\vec{x}, \omega\right)
\ .
\nonumber
\end{align}

\vspace{0.5in}

After substitution, all angular parameters are expressed in terms of spherical harmonics, but they still depend on the continuous angular variable $\omega$. As a next step, we project each term of the RTE into spherical harmonics, using the projection operator $\mathcal{P}$. This produces a single equation for each $l,m$-pair. The $P_N$-equations therefore can be written as:
\begin{align}
\mathcal{P}^{l,m}\mathcal{T}\left(\widehat{L}\right)
=
-\mathcal{P}^{l,m}\mathcal{C}\left(\widehat{L}\right) 
+\mathcal{P}^{l,m}\mathcal{S}\left(\widehat{L}\right)
+\mathcal{P}^{l,m}\left(Q\right)
\ .
\label{eq:pn_operator_notation}
\end{align}

Once the $P_N$-equations have been found, the spatial variable $\vec{x}$ is discretized using a finite difference (FD) voxel grid (using central differences for differential operators).

Following this discretization, the radiance field $L$, is represented as a set of SH coefficients per voxel. Flattening these over all voxels into a single vector, gives the solution vector $\vec{u}$. The RHS vector $\vec{Q}$ is produced similarly. The projected operators can be expressed as linear transformations, which can be collapsed into a single coefficient matrix $A$ (see Figure~\ref{fig:matrix_layout}):
\begin{align}
(T+C-S)\vec{u} = A\vec{u} = \vec{Q}
\ .
\end{align}

@@

So far, we have only given the $P_N$-equations in high-level operator notation (Equation~\ref{eq:pn_operator_notation}). Carrying out the full derivation creates large, unwieldy equations and requires a string of expansions, manipulations and applications of identities. These are challenging and error-prone if done by hand. We \rev{therefore used a computer algebra representation}, which allowed us to derive and discretize the $P_N$-equation in a semi-automatic fashion.
